Demo for a Quality of Service networking with Software Defined Network.

### 2017-03-14 report/paper.tex
added conclusions. need review. doing this evening

%\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Quality of Service with Software Defined Network.}

@@

\begin{abstract}
Quality of Service (QoS) in existing network architectures is an ongoing problem.
A lot of researchers and industries tried to solve this problem and found different solutions but some of them are too
expensive other failed or were not implemented.
Software Defined Networking (SDN) paradigm has emerged in response to traditional networking architecture's limitations.
The main advantage of SDN is a centralized global network view. This particular feature pushed researchers to keep on improving QoS.

In this report we will focus on two different aspects; all problems we faced during our attempt to create a based SDN application able to
support a QoS service and the results we achieved.

OpenFlow standard and python POX controller, which is largely spread, were used as instruments to make this experience possible,
despite the limit of using a new type of tecnology still in the research phase.
%corretto
\end{abstract}

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is a way to manage networks that separates the control from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brain” of the network.
The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it isn’t the only SDN standard. 
Centralized, programmable SDN environments can easily adjust to the rapidly changing business'needs. SDN can reduce costs and restrict
wasteful provisioning, as well as provide flexibility and innovation for networks.
%corretto
\subsection{OpenFlow}
  During 2006 ,thanks to M. Casado, a new network architecture was developed, called Ethane.
  On this conncept was born, during 2011, the Open Network Foundation, a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 was released on 2009 but the last standardized version v1.5.1 was approved only on march 2015 \cite{ONF_report}.
  %corrects
  %todo spostare l'immagine correttamente
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.58]{images/of.jpg}
 \caption{\emph{OpenFlow}}
 \label{fig:topo}
\end{figure}
  \\
  \textbf{Openflow 1.0 messages}

Openflow defines a number of messages between a controller and different switches. Now there will be presented some messages we exploited:\cite{openflow}
\begin{itemize}

\item \emph{Hello} \newline This message is sent both from the controller and the switch who is looking for a controller after it
	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
\item \emph{FeatureReq} \newline This is the message that the controller sends to a switch to get to know its abilities. The switch
	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
	With the featureRes the controller will know how many ports a switch has and their properties (e.g. if they are connected, link speed,...), the flow table capabilities and the supported statistics.
\item \emph{PortStatus} \newline this message is sent from a switch to the controller, to notify that some ports has changed their status.
		It's useful to check if a port has someone connected to it.
\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of 
statistic the controller wants to know.
		The switch then will answer with the Stats reply, if it supports that type of stat. If not, The switch will send a Bad_Request
		message.
\item \emph{PacketIn} \newline This message is sent from the switch to the controller everytime a packet arrives without matching any rule.
The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
		a flow rule for packets like this using a FlowMod.
\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message,
to some ports.
\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch.This is the message that allows routing.
\\
\\
\end{itemize}

\textbf{Forwarding table:}
\begin{itemize}
 \item \texttt{match fields:} fields that compare packets in order to look fo sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze. Contengono la porta d'ingresso, l'intestazione del pacchetto,
 e altri campi opzionali;
 \item \texttt{priority:} precedenza nel caso di voci con pi\`u corrispondenze;
 \item \texttt{counter:} incrementato ogni qualvolta che viene riscontrata la corrispondenza con un pacchetto;
 \item \texttt{istructions:} istruzioni da aggiungere all'\emph{Action-Set} o modifiche al flusso della pipeline;
 \item \texttt{timeouts:} massimo tempo per cui lo switch tiene in memoria questa voce, altre il quale la voce viene scartata;
 \item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
 \item \texttt{flags:} utilizzate per differenziare la gestione della voce;
 \\
 \\
\end{itemize}

@@

\subsection{Graph and Network structures}
  In order to achieve the realtime adaptability and statistics storage we have develop ad-hoc structure.
  We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent
  resplectively switch or host element.
  This topological structure is updated realtime thanks to connection event and link discover offered by various library.
  ( mettere i vari eventi che sfrutto )
  The link besides have many parameters in witch we can save network performance. To update every link parameter we have to creat ad hoc function
  because there isn't some function constructed yet. ( sistemare frase) .
  This type of structure allow us to run Dijkstra algorithm that return the minimum path ( in base a vari parametri ) on a nlog n time ( verificare).
  (aggiungere qualcosa e correggere)

  \begin{figure}[!h]
 \centering
 \includegraphics[scale=0.30]{images/topology.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
  \end{figure}

  \#TODO matteo

@@

	\#TODO andrea
	The first data that was collected was the actual link capacity, which depends both on the cable used to connect
	different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
	openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
	the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
	\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
	as it gives the most useful about the link state. This information, moreover, is the one that non-SDN switches use to choose the \textit{metric}
	parameter. The simulation with Mininet showed us that this type of information can be used, but the Routeboards didn't support it.
	As a matter of fact, we weren't able to exploit this capacity in our network realization.
		\\
		\\
  \subsubsection{Throughput}
	\#TODO andrea
	Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
	We exploited the information given by the switches with the match structure: this means that we used the throughput for the TCP traffic,
	the one for the UDP and for each of them we could know toward which IP the traffic is directed or is from. This was exploited with mininet
	simulations, but not in the set up in the laboratory. With the latter, we used the statistics as an indicator of the bitrate of a link
	and divided the host to categories based on it.
\\
\subsubsection{Packet error rate}

The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
heavy traffic was directed to links with highest packet error rate.  As the switches exchanges a packet before recognizing most of the error, both because the
phisical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
those are indirect measures themself and the error derived from the calculations could grow up.
\#TODO andrea
\\
\subsection{Route changing} %cambio rotte

\#TODO andrea
For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
route to reach the destination and viceversa. The traffic follow the same path in both the directions and depends only on the type of traffic:
we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
Two threshold were chosen based on the traffic analysis described on Sec. \ref{subsec:traf} which are 200 bytes/s for the gaming traffic and one megabyte/s
when the node is considered in heavy traffic. Those two threshold were made as they are average values and considering that our network is made of few really fast link,
so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, wherease it's easy to reach the high level of traffic load between two machines.
The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
load in the network will be heavy and last longer.

\subsection{ARP solution}

During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
it higher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
this is necessary.

@@

\subsection{Ip switching in pox}

During our tests, both with Mininet and in the lab, we found out that the core of POX, when a PacketIn event occoures, switches the source and the destination addresses,
both MAC and IP, if one of the hosts is already known. We weren't able to understand in which point of the code of POX  and why it happens.
% One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
% switches the source and the destination IP and MAC address with the source one.
The problem was partially solved, checking out the
parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
If the two host are in the same switch, the problem is solved using a controlled packet flooding, described in section \ref{sec:connectivity}

\section{Conclusions}\label{sec:conclusion}
Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
ehi
\#TODO

>>>>

\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Demo for a Quality of Service networking with Software Defined Network.}

@@

\begin{abstract}
Quality of Service(QoS) in existing network architectures is an ongoing problem.
A lot of researcher and industries try to solve this problem with many solution but a lot of them are too
expensive other failed or are not implemented.
Software Defined Networking(SDN) paradigm has emerged in response to limitations of traditional networking architectures.
The main advantages of SDN are centralized global network view. This particular features have got attention of researchers to improve QoS.
In questo report mostreremo quali sono stati i principali risutati raggiunti e i problemi incontrati nella realizzazione di
un applicativo SDN in grado di supportare un servizio di tipo QoS.
Lo standard OpenFlow e il controller python POX largamente diffuso sono stati gli strumenti che hanno reso possibile la
realizzazione di quest'esperienza pur mostrando i limiti dovuti all'utilizzo di una tecnologia nuova e in fase di ricerca.
%translate
\end{abstract}

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is way to manage networks that separates the control plane from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brains” of the network.
The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, however, it isn’t the only SDN standard, but it is widely accepted.
Centralized, programmable SDN environments can easily adjust to the rapidly changing needs of businesses. SDN can lower costs and limit
wasteful provisioning, as well as provide flexibility and innovation for networks.
\#TODO da completare
  \subsection{OpenFlow}
  During 2006 thanks to M. Casado was developed Ethane a new network architecture on this conncept during 2011 was bord the Open Network Foundatin.
  This is a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 release on 2009 but the last standardized version is v1.5.1 approved on march 2015 \cite{ONF_report}.
  \#TODO da sistemare
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.58]{images/of.jpg}
 \caption{\emph{OpenFlow}}
 \label{fig:topo}
\end{figure}
  \\
  \textbf{Openflow 1.0 messages}
Openflow defines a number of messages between a controller and a switch. Now there will be presented some messages we exploited:\cite{openflow}
\begin{itemize}

\item \emph{Hello} \newline This message is sent both from the controller and the switch who is looking for a controller after it
	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
\item \emph{FeatureReq} \newline this is the message that the controller sends to a switch to know its abilities. The switch
	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
	With the featureRes the controller will know how many ports a switch has and their properties (e.g. if they are connected, link speed,...), the flow table capabilities and the supported statistics.
\item \emph{PortStatus} \newline this message is sent from a switch to the controller, to notify that some ports has changed their status.
		It's useful to check if a port has someone connected
\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of statistic the controller want to know.
		The switch then will answer with the Stats reply, if it supports that type of stat. It will send a Bad_Request message if not.
\item \emph{PacketIn} \newline This message is sent from the switch to the controller whenever a packet arrives and it doesn't match
		any rule. The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
		a flow rule for packets like this using a FlowMod.
\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message, to some
		port(s).
\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch. Is the message that allows routing.
\\
\\
\end{itemize}

\textbf{Forwarding table:}
\begin{itemize}
 \item \texttt{match fields:} sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze. Contengono la porta d'ingresso, l'intestazione del pacchetto,
 e altri campi opzionali;
 \item \texttt{priority:} precedenza nel caso di voci con pi\`u corrispondenze;
 \item \texttt{counter:} incrementato ogni qualvolta che viene riscontrata la corrispondenza con un pacchetto;
 \item \texttt{istructions:} istruzioni da aggiungere all'\emph{Action-Set} o modifiche al flusso della pipeline;
 \item \texttt{timeouts:} massimo tempo per cui lo switch tiene in memoria questa voce, altre il quale la voce viene scartata;
 \item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
 \item \texttt{flags:} utilizzate per differenziare la gestione della voce;
 \\
 \\
\end{itemize}

@@

\subsection{Graph and Network structures}
  In order to achieve the realtime adaptability and statistics storage we have develop ad-hoc structure.
  We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent
  resplectively switch or host element.
  This topological structure is updated realtime thanks to connection event and link discover offered by various library.
  ( mettere i vari eventi che sfrutto )
  The link besides have many parameters in witch we can save network performance. To update every link parameter we have to creat ad hoc function
  because there isn't some function constructed yet. ( sistemare frase) .
  This type of structure allow us to run Dijkstra algorithm that return the minimum path ( in base a vari parametri ) on a nlog n time ( verificare).
  (aggiungere qualcosa e correggere)
  \#TODO matteo

@@

	The first data that was collected was the actual link capacity, which depends both on the cable used to connect
	different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
	openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
	the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
	\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
	as it gives the most useful about the link state. This information, moreover, is the one that non-SDN switches use to choose the \textit{metric}
	parameter. The simulation with Mininet showed us that this type of information can be used, but the Routeboards didn't support it.
	As a matter of fact, we weren't able to exploit this capacity in our network realization.
		\\
		\\
  \subsubsection{Throughput}
	Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
	We exploited the information given by the switches with the match structure: this means that we used the throughput for the TCP traffic,
	the one for the UDP and for each of them we could know toward which IP the traffic is directed or is from. This was exploited with mininet
	simulations, but not in the set up in the laboratory. With the latter, we used the statistics as an indicator of the bitrate of a link
	and divided the host to categories based on it.
\\
\subsubsection{Packet error rate}
The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
heavy traffic was directed to links with highest packet error rate.  As the switches exchanges a packet before recognizing most of the error, both because the
phisical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
those are indirect measures themself and the error derived from the calculations could grow up.
\#TODO andrea
\\
\subsection{Route changing} %cambio rotte
For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
route to reach the destination and viceversa. The traffic follow the same path in both the directions and depends only on the type of traffic:
we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
Two threshold were chosen based on the traffic analysis described on Sec. \ref{subsec:traf} which are 200 bytes/s for the gaming traffic and one megabyte/s
when the node is considered in heavy traffic. Those two threshold were made as they are average values and considering that our network is made of few really fast link,
so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, wherease it's easy to reach the high level of traffic load between two machines.
The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
load in the network will be heavy and last longer.

\subsection{ARP solution}
During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
it higher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
this is necessary.

@@

\subsection{Ip switching in pox}
During our tests, both with Mininet and in the lab, we found out that the core of POX, when a PacketIn event occoures, switches the source and the destination addresses,
both MAC and IP, if one of the hosts is already known. We weren't able to understand in which point of the code of POX  and why it happens.
% One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
% switches the source and the destination IP and MAC address with the source one.
The problem was partially solved, checking out the
parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
If the two host are in the same switch, the problem is solved using a controlled packet flooding, described in section \ref{sec:connectivity}

\section{Conclusions}\label{sec:conclusion}
% Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
% ehi
For this course project, the report shows how the demo was made, in particular what problem we encountered
and what we exploited, showing what protocols could be unused in the future. We showed a possible simple data structure
that can allow the controller to be able to know the network, its topology and which route to set, based on the traffic.
Morover, we showed how to exploit the potentialities of the SDN network, giving particular importance on the
statistics and the path switching capabilities.
The multi-graph solution allowed us to connect the network and, based on its condition, choose which weight to apply
in order to give a Quality of Service based network. With some boards that supports the newer version of Openflow
(i.e. 1.4, the latest), the code can be ported, so it can be used with the Ryu core controller and more statistic can
be exploited. Moreover, the controller can set more strictive rules for the flow control and can get more informations from the
switches.
